Chemical tagging can work: \\ Identification of stellar phase-space structures \\ purely by chemical-abundance similarity

### 2016-02-21 papers/kmeans/response_to_referee.txt

Dear Editor:

Please find below the response to our (very helpful) referee report on
manuscript ApJ/ApJ101852.  We have taken all of the referees
suggestions, and made some other changes from comments we have
received from colleagues.

In what follows, the Referee comments are indented and our responses
are not.  In the revised manuscript, we have made the changes in bold
text so they are apparent to the editor and referee.

     This paper presents new element abundance results using The
     Cannon, a novel, powerful method for spectroscopic abundance
     measurements. The results presented in this paper are new and
     very much relevant to the many large scale observing projects
     that are currently in operation.

     I highlight some comments for the authors to address:

     Abstract: The last sentence is not true, as the authors note
     later in the text the first blind search purely in abundance
     space was carried out by Mitschang et al 2014.

HOGG

     Introduction: The first paragraph is missing a reference or two,
     to support claim that most stars end up in different parts of the
     Galaxy. Or this is something that is assumed and yet to be
     tested?

HOGG

     There are some missing references to the original demonstrations
     of chemical tagging of unbound groups, namely the Wolf 630 group
     and HR 1614 group. Those were key to building up the case for
     this "tagging" research.

HOGG

     The text needs to highlight that the training set is critical -
     Cannon doesn't know physics. It needs to be told what is "true"
     and that the subsequent results will be limited to the believed
     level of truth. While it should be specified that tagging per se
     doesn't require absolute abundances, but the subsequent
     interpretation of the chemical accretion history will require
     "true" abundances - not just precision. Also some clarification
     about the Cannon's performance at the boundaries of the training
     set parameters would have been good to see.

HOGG

     Section 2:

     A reference for the famously demonstrated correlations in
     globular clusters is needed.

HOGG

     Section 3:

     The text doesn't clarify why only k-means was chosen to do the
     'tagging' given the issues discussed, was it simply easy and
     fast? Fine as the focus of the paper is the precision of the
     abundance results with Cannon, not the fidelity of group finding.

HOGG

     I find the group in Fig 6 dubious. Its at the very edge of the
     metallicity scale, which raises again how the Cannon performs at
     the boundaries of the training parameters. Could this not simply
     be metal-poor Halo stars, tagged as metal-poor Halo stars,
     nothing more?

HOGG

-----
Other changes:

In addition to all the above, we made the following additional
changes:

HOGG:  We added authors!

HOGG:  We added citations!

HOGG:  Adjusted what we say about Ting et al?

HOGG:  Go through github issues and emails.

### 2016-02-27 papers/kmeans/response_to_referee.txt

HOGG

-----
Other changes:

In addition to all the above, we made the following additional
changes:

Because of a large number of valuable contributions both prior to and
subsequent to our submission of this paper, we expanded the author
list.  There are now several additional new authors, most of whom
contributed by building the APOGEE project (and hence the input data).

We added a few citations: Eisenstein 2011 (SDSS-III), Holtzman 2015
(APOGEE DR12) and ...HOGG

@@

HOGG:  Adjusted what we say about Ting et al?

HOGG:  Go through github issues and emails.

We also added a few names to the acknowledgements for providing minor
comments.

### 2016-02-27 papers/kmeans/kmeans.tex
citations, authors

\title{Chemical tagging can work: \\
       Identification of stellar phase-space structures \\
       purely by chemical-abundance similarity}
\author{David~W.~Hogg\altaffilmark{1,2,3,4},
        Andrew~R.~Casey\altaffilmark{5},
        Melissa~Ness\altaffilmark{4},
        Hans-Walter~Rix\altaffilmark{4},
    and~Daniel~Foreman-Mackey\altaffilmark{6,7}}
\altaffiltext{1}{Simons Center for Data Analysis, 160 Fifth Avenue, 7th floor, New York, NY 10010, USA}
\altaffiltext{2}{Center for Cosmology and Particle Physics, Department of Physics,
  New York University, 4 Washington Pl., room 424, New York, NY 10003, USA}
\altaffiltext{3}{Center for Data Science, New York University, 726 Broadway, 7th floor, New York, NY 10003, USA}
\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
\altaffiltext{5}{Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge CB3~0HA, UK}
\altaffiltext{6}{Sagan Fellow}
\altaffiltext{7}{Department of Astronomy, University of Washington, Box 351580, Seattle, WA 98195, USA}
\email{david.hogg@nyu.edu}

@@

In what follows, we are going to use \apogee\ \acronym{DR12} data \citep{dr12}, in
which we can re-derive 15 element abundances (C, N, O, Na, Mg, Al, Si,
S, K, Ca, Ti, V, Mn, Fe, Ni), using \thecannon\ \citep{thecannon}.
The \apogee\ data set covers a huge radial extent and---because the
data are taken in the infrared---explores the thin disk.
However, it has the disadvantage that its spatial coverage is incomplete
(that is, limited pointings produce a fractured sky map),
which makes it hard to see, within the data set, linear or
extended stellar structures.
In many ways, \galah\ will deliver improvements, both because it will
have more abundances (possibly 29) and contiguous sky coverage.

@@

Our analysis draws on the spectra of \totalnumber\ giant stars ($\logg
< 3.9$) from \apogee\ \acronym{DR12} \citep{dr12}, with no warning flags
set in the \apogee\ \aspcap\ \citep{aspcap} pipeline
reductions.
We re-analyze these \emph{spectra} using \thecannon, because it can
deliver stellar abundance labels of higher precision, especially for
stars of $S/N \le 150$.
The details about how we select, reduce, and analyze the \apogee\ data
are given in full detail in the companion papers \citep{casey16,
  ness16}, and we only summarize briefly here:
We re-derived 17 labels ($\teff$, $\logg$, and 15 abundances
referenced to H) from the \apogee\ \acronym{DR12} spectra. 
For the training step of \thecannon, 
12,681 red-giant stars with spectral $S/N \ge 200$ were used.
For this small fraction of the sample with the highest $S/N$, the
labels provided by \aspcap\ provide consistent, relatively low-scatter, sensible
abundance-space measurements.
The training step fixes a spectral model
that predicts the normalized spectrum as a quadratic function of all
labels ($\sim 1.5\times10^6$ model coefficients). In the second stage---the test step---each unlabelled star is
used to establish a single-star likelihood function for its labels,
holding the spectral model coefficients (the parameters of the model) fixed.
\thecannon\ finds the labels that minimize the single-star likelihood
function for each test-set star.
This optimization is not convex, but it is trivially parallel,
and therefore fast. The test
step for 150,000 spectra takes less than 30~minutes on a small 
research cluster in Cambridge.

@@

\acknowledgements
It is a pleasure to thank
  Anna~Y.~Q.~Ho (Caltech),
  Jeremy~Magland (\acronym{SCDA}), and
  the Blanton--Hogg group meeting
for valuable comments and discussions.
This project was funded in part by
  the \acronym{NSF} (grants \acronym{IIS-1124794}, \acronym{AST-1517237}),
  \acronym{NASA} (grant \acronym{NNX12AI50G}),
  the Moore-Sloan Data Science Environment at \acronym{NYU}, and
  the European Research Council under the
  European Union's Seventh Framework Programme (FP~7)
  \acronym{ERC} Grant Agreement n.~\acronym{[320360, 321035]}).
This research made use of the \acronym{NASA} \project{Astrophysics Data System}.

@@

@@

>>>>

\title{Chemical tagging can work: \\
       Identification of stellar phase-space structures \\
       purely by chemical-abundance similarity}
\author{David~W.~Hogg\altaffilmark{1,2,3,4},
        Andrew~R.~Casey\altaffilmark{5},
        Melissa~Ness\altaffilmark{4},
        Hans-Walter~Rix\altaffilmark{4},
        Daniel~Foreman-Mackey\altaffilmark{6,7},
        Anna~Y.~Q.~Ho\altaffilmark{8},
        Jon~A.~Holtzman\altaffilmark{9},
        David~L.~Nidever\altaffilmark{10}}
\altaffiltext{1}{Simons Center for Data Analysis, 160 Fifth Avenue, 7th floor, New York, NY 10010, USA}
\altaffiltext{2}{Center for Cosmology and Particle Physics, Department of Physics,
  New York University, 4 Washington Pl., room 424, New York, NY 10003, USA}
\altaffiltext{3}{Center for Data Science, New York University, 726 Broadway, 7th floor, New York, NY 10003, USA}
\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
\altaffiltext{5}{Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge CB3~0HA, UK}
\altaffiltext{6}{Sagan Fellow}
\altaffiltext{7}{Department of Astronomy, University of Washington, Box 351580, Seattle, WA 98195, USA}
\altaffiltext{8}{Astronomy Department, California Institute of Technology, MC 249-17, 1200 East California Blvd, Pasadena, CA 91125, USA}
\altaffiltext{9}{New Mexico State University, Las Cruces, NM 88003, USA}
\altaffiltext{10}{Steward Observatory, 933 North Cherry Ave, Tuscon, AZ 85719, USA}
\email{david.hogg@nyu.edu}

@@

In what follows, we are going to use \apogee\ \acronym{DR12}
data \citep{apogeedr12, dr12}
from \sdssiii\ \citep{sdssiii}, in
which we can re-derive 15 element abundances (C, N, O, Na, Mg, Al, Si,
S, K, Ca, Ti, V, Mn, Fe, Ni), using \thecannon\ \citep{thecannon}.
The \apogee\ data set covers a huge radial extent and---because the
data are taken in the infrared---explores the thin disk.
However, it has the disadvantage that its spatial coverage is incomplete
(that is, limited pointings produce a fractured sky map),
which makes it hard to see, within the data set, linear or
extended stellar structures.
In many ways, \galah\ will deliver improvements, both because it will
have more abundances (possibly 29) and contiguous sky coverage.

@@

Our analysis draws on the spectra of \totalnumber\ giant stars ($\logg
< 3.9$) from \apogee\ \acronym{DR12} \citep{apogeedr12}, with no warning flags
set in the \apogee\ \aspcap\ \citep{aspcap} pipeline
reductions.
We re-analyze these \emph{spectra} using \thecannon, because it can
deliver stellar abundance labels of higher precision, especially for
stars of $S/N \le 150$.
The details about how we select, reduce, and analyze the \apogee\ data
are given in full detail in the companion papers \citep{casey16,
  ness16}, and we only summarize briefly here:
We re-derived 17 labels ($\teff$, $\logg$, and 15 abundances
referenced to H) from the \apogee\ \acronym{DR12} spectra. 
For the training step of \thecannon, 
12,681 red-giant stars with spectral $S/N \ge 200$ were used.
For this small fraction of the sample with the highest $S/N$, the
labels provided by \aspcap\ provide consistent, relatively low-scatter, sensible
abundance-space measurements.
The training step fixes a spectral model
that predicts the normalized spectrum as a quadratic function of all
labels ($\sim 1.5\times10^6$ model coefficients). In the second stage---the test step---each unlabelled star is
used to establish a single-star likelihood function for its labels,
holding the spectral model coefficients (the parameters of the model) fixed.
\thecannon\ finds the labels that minimize the single-star likelihood
function for each test-set star.
This optimization is not convex, but it is trivially parallel,
and therefore fast. The test
step for 150,000 spectra takes less than 30~minutes on a small 
research cluster in Cambridge.

@@

\acknowledgements
It is a pleasure to thank
  Jeremy~Magland (\acronym{SCDA}),
  Don Schneider (PSU), and
  the Blanton--Hogg group meeting
for valuable comments and discussions.
This project was funded in part by
  the \acronym{NSF} (grants \acronym{IIS-1124794}, \acronym{AST-1517237}),
  \acronym{NASA} (grant \acronym{NNX12AI50G}),
  the Moore-Sloan Data Science Environment at \acronym{NYU}, and
  the European Research Council under the
  European Union's Seventh Framework Programme (FP~7)
  \acronym{ERC} Grant Agreement n.~\acronym{[320360, 321035]}).
This research made use of the \acronym{NASA} \project{Astrophysics Data System}.

@@

\bibitem[Eisenstein \etal(2011)]{sdssiii}
Eisenstein, D.~J., Weinberg, D.~H., Agol, E., \etal, 2011, \aj, 142, 72 

@@

\bibitem[Holtzman \etal(2015)]{apogeedr12}
Holtzman, J.~A., Shetrone, M., Johnson, J.~A., \etal, 2015, \aj, 150, 148
